Meeting notes

### 2023-04-03 TeX/draft.tex
Updates from Overleaf

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}%[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem*{proposition*}{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem*{corollary*}{Corollary}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{algorithm}[theorem]{Algorithm}

%\numberwithin{equation}{section}

@@

% Nongross real and imaginary parts
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\Q}{\bQ}
\newcommand{\Z}{\bZ}
\newcommand{\filip}[1]{{\color{cyan} \textsf{$\sun\sun\sun$ Filip: [#1]}}}
\newcommand{\davidnote}[1]{\marginpar{\footnotesize{\textcolor{Plum}{David: #1}}}}

@@

>>>>

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}%[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem*{proposition*}{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem*{corollary*}{Corollary}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{algorithm}[theorem]{Algorithm}
\DeclareMathOperator{\gon}{gon}
%\numberwithin{equation}{section}

@@

% Nongross real and imaginary parts
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\Q}{\bQ}
\newcommand{\Z}{\bZ}
\newcommand{\PP}{\mathbb P}
\newcommand{\filip}[1]{{\color{cyan} \textsf{$\sun\sun\sun$ Filip: [#1]}}}
\newcommand{\davidnote}[1]{\marginpar{\footnotesize{\textcolor{Plum}{David: #1}}}}

@@

\section{Genus 0 adelic images do not produce isolated points}

Let $E/ \Q$ be an elliptic curve and $G\in \GL_2(\hat \Z)$ its adelic image. By $G(n)$ denote the image of its mod $n$ representation. Denote by $B_1(n)$ the subgroup of $\GL_2(\Z /n \Z)$ consisting of the upper triangular matrices with a $1$ in the upper left entry. Note that $X_{B_1(n)}=X_1(n)$. We say that a congruence group $\Gamma$ is of genus $g$ if $X_\Gamma$ is of genus $g$. We say that a point $P$ corresponds to an elliptic curve $E$ if $j(P)=j(E)$.

\begin{proposition}
Let $E/\Q$ be an elliptic curve with adelic image of genus $0$. Then there does not exist an $n$ such that a point corresponding to $E$ on $X_1(n)$ is sporadic.
\end{proposition}
\begin{proof}
Since $G$ is by assumption of genus $0$, it follows that so is $G(n)$. Let $P$ be a $\Q$-rational point on $X_{G(n)}$ corresponding to $E$.

%Let $B$ be the group generated by $G(n)$ and a conjugate of $B_1(n)$ in $\GL_2(\Z/n\Z)$; this is again a group of genus $0$ on which the image $P'$ of $P$ under the map $X_{G(n)}\rightarro X_B$ is $\Q$-rational. The inclusion of subgroups $B\subset B_1(n)$ induces a map $X_1(n)\rightarrow  X_B$.

The minimal degree of a point on $X_1(n)$ corresponding to $E$ is the minimal index of $B:=c(B_1(n))\cap G(n)$ in $G(n)$, where $c$ denotes conjugation in $\GL_2(\Z /n \Z)$. 
Call this (minimal) index $d$ and let $f:X_{B}\rightarrow X_{G(n)}$ be the corresponding map of modular curves. By construction, all the points in $f^{-1}(P)$ are of degree $d$ and obviously $\gon_\Q X_{B}\leq d.$ Let $P'\in f^{-1}(P) \subseteq X_{B}$ be a point corresponding to $E$, and let $g:X_{B}\rightarrow X_1(n)$ be the map induced by the inclusion $B_1(n) \subseteq B$. Then by the argument above we have that $g(P')$ is of degree $d$ (as this is the minimal degree of a point on $X_1(n)$ corresponding to $E$), and on the other hand we have that $\gon_\Q X_1(n)\leq d$ by \cite[Proposition A.1 (vii)]{Poonen:gonality}.
\end{proof}

### 2023-04-03 TeX/notes/meeting_notes_29Mar2023.tex
Updates from Overleaf

\usepackage{minted}
\usemintedstyle{lovelace}
\definecolor{mintbf}{gray}{0.95}
\setminted{bgcolor=mintbf}

@@

\begin{minted}[mathescape, linenos, samepage]{c}
/* First half of NotIsolated function, which determines if j
 * can't give isolated points. */
NotIsolated:=function(a, j);             // #1
    E:=EllipticCurve(a);
    G,n,S:=FindOpenImage(path, E);       // #2
    G0:=ReducedLevel(G);                 // #3 (Things to be proved!)
    G0t := sub<
      GL(2,Integers(#BaseRing(G0))) | [Transpose(g):g in Generators(G0)]
    >;                                   // #4
    k:=#BaseRing(G0t);
    good:=[];                            // "not isolated"
    bad:=[];                             // "potentially isolated"
    for b in Divisors(k) do
        if b gt 12 then                  // #5
            ww:=MinDegreeOfPoint(ChangeRing(G0t,Integers(b)));
                                         // #6
            // if the min degree >= g+1 then the dimension of the Reimann-Roch
            // space associated to the point of min degree is at least 2 hence it
            // is not sporadic.
            if  ww ge (Genus(Gamma1(b))+1) then
                Append(~good,<b, ww>);   // #7
            else
                // otherwise we need to check if gonality/degree reduction
                // arguments can be used
                Append(~bad,<b, ww>);
            end if;
        end if;
    end for;
\end{minted}

@@

\begin{minted}[mathescape, linenos, firstnumber=last, samepage]{c}
    /* continuing from before */
    // Refuting levels and degrees based on the level reduction theorem of BELOV.
    // If j is a sporadic point of degree d in level m then it becomes a point of
    // d/deg(f) in level n where f:X1(m)-->X1(n) is the natural projection map.
    remove:={};                          // #8 START
    for x in bad do
        for y in good do
            if IsDivisibleBy(x[1],y[1]) then
                b:=x[1] div y[1];
                deg:=b^2*&*[Rationals()
                              | 1-1/p^2 : p in PrimeDivisors(x[1])
                              | p notin PrimeDivisors(y[1]) ];
                if deg eq x[2] div y[2] then Include(~remove,x); end if;
            end if;
        end for;
    end for;                             // #8 END

    // Now we check using gonality arguments if the levels, degrees in the set
    // bad but not in remove can be handled using gonality arguments.
    bad := SequenceToSet(bad);
    if bad ne remove then                // #9
        supbad := bad diff remove;
        return [*j, a, false, supbad*]; //return j-invariant we already have
    end if;
    return [*j, a, true*];
end function;
\end{minted}

@@

\begin{minted}[mathescape, linenos, samepage]{c}
/* NOTE: We explicitly check those primes dividing m when G doesn't surject to
 * GL2FP. */
NonSurjectivePrimes:=function(G)                     // #1
    m:=Modulus(BaseRing(G));
    return [p:p in PrimeFactors(m)|#ChangeRing(G,GF(p)) ne #GL(2,GF(p))];
end function;

// Level reduction from the output of Zywina's algorithm
ReducedLevel:=function(G)
    m:=Modulus(BaseRing(G));
    NS:=Set(NonSurjectivePrimes(G));
    sE:={2,3} join NS;                    // always consider 2 and 3
    m0:=&*[p^Valuation(m,p):p in sE];
    G:=ChangeRing(G,Integers(m0));
    for p in PrimeFactors(m0) do                     // #2
        while Valuation(m0,p) gt 1                   // #3
                and #G/#ChangeRing(G,Integers(m0 div p)) eq p^4 do
            m0:=m0 div p;
            G:=ChangeRing(G,Integers(m0));
        end while;
        if not p in NS and Valuation(m0,p) eq 1      // #4
               and #G/#ChangeRing(G,Integers(m0 div p)) eq #GL(2,GF(p)) then
            m0:=m0 div p;
            G:=ChangeRing(G,Integers(m0));
        end if;
    end for;
    return G,m0;
end function;
\end{minted}

@@

\begin{minted}[mathescape, linenos, samepage]{c}
VectorOrder:=function(v)
    m:=#Parent(v[1]);
    g:=GCD([m, Integers()!v[2], Integers()!v[1]]);
    return m div g;
end function;

//compute min degree of a point i.e. the min degree of the field of definition
//of a point expressed as a 1 x 2  vector.
MinDegreeOfPoint:=function(G)
    m:=Modulus(BaseRing(G));
    H:=sub<GL(2,Integers(m))|G,-G!1>;           // #1
    orb:=Orbits(H);                             // #2
    sorb:=Sort(orb,func<o1,o2|#o1-#o2>);        // #3
    s2:=[x: x in sorb| VectorOrder(x[1]) eq m]; // #4
    return (#s2[1]) div 2;                      // #5
end function;
\end{minted}

>>>>

% Don't use minted, b/c overleaf
% \usepackage{minted}
% \usemintedstyle{lovelace}
% \definecolor{mintbf}{gray}{0.95}
% \setminted{bgcolor=mintbf}

@@

\begin{verbatim}
/* First half of NotIsolated function, which determines if j
 * can't give isolated points. */
NotIsolated:=function(a, j);             // #1
    E:=EllipticCurve(a);
    G,n,S:=FindOpenImage(path, E);       // #2
    G0:=ReducedLevel(G);                 // #3 (Things to be proved!)
    G0t := sub<
      GL(2,Integers(#BaseRing(G0))) | [Transpose(g):g in Generators(G0)]
    >;                                   // #4
    k:=#BaseRing(G0t);
    good:=[];                            // "not isolated"
    bad:=[];                             // "potentially isolated"
    for b in Divisors(k) do
        if b gt 12 then                  // #5
            ww:=MinDegreeOfPoint(ChangeRing(G0t,Integers(b)));
                                         // #6
            // if the min degree >= g+1 then the dimension of the Reimann-Roch
            // space associated to the point of min degree is at least 2 hence it
            // is not sporadic.
            if  ww ge (Genus(Gamma1(b))+1) then
                Append(~good,<b, ww>);   // #7
            else
                // otherwise we need to check if gonality/degree reduction
                // arguments can be used
                Append(~bad,<b, ww>);
            end if;
        end if;
    end for;
\end{verbatim}

@@

\begin{verbatim}
    /* continuing from before */
    // Refuting levels and degrees based on the level reduction theorem of BELOV.
    // If j is a sporadic point of degree d in level m then it becomes a point of
    // d/deg(f) in level n where f:X1(m)-->X1(n) is the natural projection map.
    remove:={};                          // #8 START
    for x in bad do
        for y in good do
            if IsDivisibleBy(x[1],y[1]) then
                b:=x[1] div y[1];
                deg:=b^2*&*[Rationals()
                              | 1-1/p^2 : p in PrimeDivisors(x[1])
                              | p notin PrimeDivisors(y[1]) ];
                if deg eq x[2] div y[2] then Include(~remove,x); end if;
            end if;
        end for;
    end for;                             // #8 END

    // Now we check using gonality arguments if the levels, degrees in the set
    // bad but not in remove can be handled using gonality arguments.
    bad := SequenceToSet(bad);
    if bad ne remove then                // #9
        supbad := bad diff remove;
        return [*j, a, false, supbad*]; //return j-invariant we already have
    end if;
    return [*j, a, true*];
end function;
\end{verbatim}

@@

\begin{verbatim}
/* NOTE: We explicitly check those primes dividing m when G doesn't surject to
 * GL2FP. */
NonSurjectivePrimes:=function(G)                     // #1
    m:=Modulus(BaseRing(G));
    return [p:p in PrimeFactors(m)|#ChangeRing(G,GF(p)) ne #GL(2,GF(p))];
end function;

// Level reduction from the output of Zywina's algorithm
ReducedLevel:=function(G)
    m:=Modulus(BaseRing(G));
    NS:=Set(NonSurjectivePrimes(G));
    sE:={2,3} join NS;                    // always consider 2 and 3
    m0:=&*[p^Valuation(m,p):p in sE];
    G:=ChangeRing(G,Integers(m0));
    for p in PrimeFactors(m0) do                     // #2
        while Valuation(m0,p) gt 1                   // #3
                and #G/#ChangeRing(G,Integers(m0 div p)) eq p^4 do
            m0:=m0 div p;
            G:=ChangeRing(G,Integers(m0));
        end while;
        if not p in NS and Valuation(m0,p) eq 1      // #4
               and #G/#ChangeRing(G,Integers(m0 div p)) eq #GL(2,GF(p)) then
            m0:=m0 div p;
            G:=ChangeRing(G,Integers(m0));
        end if;
    end for;
    return G,m0;
end function;
\end{verbatim}

@@

\begin{verbatim}
VectorOrder:=function(v)
    m:=#Parent(v[1]);
    g:=GCD([m, Integers()!v[2], Integers()!v[1]]);
    return m div g;
end function;

//compute min degree of a point i.e. the min degree of the field of definition
//of a point expressed as a 1 x 2  vector.
MinDegreeOfPoint:=function(G)
    m:=Modulus(BaseRing(G));
    H:=sub<GL(2,Integers(m))|G,-G!1>;           // #1
    orb:=Orbits(H);                             // #2
    sorb:=Sort(orb,func<o1,o2|#o1-#o2>);        // #3
    s2:=[x: x in sorb| VectorOrder(x[1]) eq m]; // #4
    return (#s2[1]) div 2;                      // #5
end function;
\end{verbatim}
